\chapter*{Aknowledgments}
\addcontentsline{toc}{chapter}{Aknowledgments} % Add the preface to the table of contents as a chapter

There are many people thanks to whom I write this work today. First of all I would like to thank my teachers Dr. Jose Emilio Labra Gayo and Daniel Fernández Álvarez, as well as the semantic web research group of the University of Oviedo for the trust placed in me to carry out this project.
However, this project is done as a summary of what has been my time at the university and a personal stage. That is why I would like to make a small dedication to those people who in one way or another have helped me to write this project today.
To my mother, Esther, for waiting until I was ready to leave us. To my uncle Andrés for being my guardian angel. To my grandparents, for giving me everything they had. To my coach Vic, who unwittingly has become a father to me. To Laura, the impossible girl. Ricardo, although I have never counted on you, you have always been. To Monica, for being you. To Pablo, without you I would not be who I am and I would never have finished this degree. To Álvaro, for letting me be your “manín". To Sari, for laughing at me when I needed it. To Cotito, for teaching me to accept myself. To Alejandro, for getting me out of my comfort zone. To Pablín, for being my prettiest doctor. To all my friends and family, those who I did not mention and those that I left along the way I dedicate this work to you for the moments you give me. And finally to my CAU family, thank you, really.
I hope that this work and I will not be a disappointment to any of you, really, each and every one of you are exceptional.

\begin{kaobox}[frametitle=Spanish]
	Son muchas las personas gracias a las que hoy escribo este trabajo. En primer lugar me gustaría agradecer a mis tutores Aquilino Juan Fuerte y Jose Emilio Labra Gayo, así como al grupo de investigación de web semántica de la Universidad de Oviedo y a Izertis S.A. la confianza depositada en mi para la realización de este proyecto.
	Sin embargo este proyecto se realiza como resumen de lo que ha sido mi paso por la universidad y por una etapa personal. Es por eso que me gustaría realizar una pequeña dedicatoria a aquellas personas que de una forma u otra han ayudado a que este escribiendo este proyecto hoy. A mi madre, Esther, por esperar a que estuviera preparado para marcharse. A mi tío Andrés por haber sido mi angel de la guarda. A mis abuelos, por dejarse la vida en mí. A mi entrenador Vic, quien sin quererlo se ha convertido en un padre. A Laura, la chica imposible. A Ricardo, aunque jamás he contado contigo siempre has estado. A Mónica, por ser tú. A Pablo, sin ti no sería quién soy y jamás habría terminado la carrera. A Álvaro, por dejarme ser tu manín. A Sari, por reñirme cuando lo necesitaba. A Cotito, por ser enseñarme a aceptarme. A Alejandro, por sacarme de mi zona de confort. A Pablín, por ser mi médico más cuqui. A todos mis amigos y amigas que dejé por el camino os dedico este trabajo por los momentos que me distéis. Y finalmente a mi familia del CAU, gracias, de verdad.
	Espero que este trabajo y yo no seamos una decepción para ninguno de vosotros, de verdad, todos y cada uno sois excepcionales.
\end{kaobox}